V1298 Tau with TESS: Updated Ephemerides, Radii, and Period Constraints for V1298~Tau~e

### 2021-10-22 src/ms.tex
anotha figure

\title{Updated Ephemerides of the young multi-planet system V1298 Tau from TESS}

@@

@@

\author{DFM?}

@@

%%%%%%%%%%%%%%%%%%%%
% abstract must be < 250 words
% currently at 180 words
%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
\sname is a young (20--30~Myr) solar analogue hosting four transiting exoplanets. Given its planets range between 0.5 - 0.9 $R_J$, this system provides a unique opportunity to understand the evolution of planetary radii in the same stellar environment. \sname was originally observed 6 years ago during K2 Campaign 4. The extended mission of NASA's Transiting Exoplanet Survey Satellite (\tess) includes observing the ecliptic plane. Here, we present new photometric observations of \sname from the 10-minute TESS Full-Frame Images. We use the TESS data to update the transit-timing for \allplanets as well as compare newly observed radii to those derived from the K2 light curve. We additionally catch a transit of \planete, which was a single-transit event in the previous data, allowing us to further constrain the period of this farthest known planet. \sname is the target of several ongoing and future observations, including with the James Webb Space Telescope. These updated ephemerides will be crucial for accurately recovering transit events and understanding any future Doppler tomographic or transmission spectroscopy signals.\end{abstract}

@@

The youth of \sname was initially established on the basis of its strong X-ray emission \citep{Wichmann1996} and high photospheric lithium abundance \citep{Wichmann2000}. A blind search for co-moving stars using Gaia DR1 astrometry data found \sname was co-moving with 8 other stars \citep[Group 29 in][]{Oh2017}. In a kinematic study of the Taurus star-forming region \citet{Luhman2018} used Gaia DR2 data to extended the membership of this group and derived an age of $\sim$~40~Myr. However, more recent analyses based on Gaia EDR3 astrometry suggests \sname may belong to either the D2 or D3 subgroups of Taurus, both of which have estimated ages $\lesssim$10~Myr \citep[][Gaidos et al. submitted]{Krolikowski2021}. Other studies focused specifically on the \sname system have estimated its age to be 23$\pm$4~Myr from comparison with empirical and theoretical isochrones \citep{David2019b}, or 28$\pm$4~Myr from isochrone fitting to the \citet{Luhman2018} Group 29 membership list given Gaia EDR3 data (Johnson et al. submitted). While the precise age of \sname remains uncertain, most estimates fall in the 10--40~Myr range and we adopt $t \approx$~20--30~Myr.

%An analysis of the Gaia DR1 astrometric data found that \sname was co-moving with 8 other stars \citep{Oh2017}. \citet{Luhman2018} used Gaia DR2 astrometry to expand the membership of this group, referred to as Oh's Group 29, and examine its relationship to the Taurus-Auriga star-forming region. \citet{Luhman2018} concluded that while Group 29 is indeed young ($\sim$40~Myr), the collection of stars was physically unrelated to the traditional star-forming groups within the Taurus-Auriga complex.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!ht]
\begin{center}
\includegraphics[width=0.46\textwidth,trim={0.25cm 0 0 0}]{static/TESSaperture.pdf}
\caption{The \tess\ target pixel file (TPF) overlaid with a sky image of \sname taken with the Sloan Digital Sky Survey (SDSS) i-band. \sname is shown in the white circle; nearby sources with \tess\ magnitudes $< 14$ are marked with orange x's. There are two bight nearby sources to \sname. However, selecting the one pixel aperture (white square) encapsulates only \sname and therefore contamination from the nearby sources is negligible. \href{https://github.com/afeinstein20/v1298tau_tess/blob/main/src/figures/tpf.py}{\github}} 
\label{fig:tpf}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

During its Extended Mission Cycle 4, \tess\ is re-observing many of the previous \textit{K2} fields. \sname (TIC 15756231) was observed by \tess\ in Sectors 43 (UT 16 Sep 2021 to UT 12 Oct 2021) at 10-minute cadence within the Full-Frame Images (FFIs). We used the \texttt{tica} \citep{fausnaugh20} software to download calibrated FFIs each orbit for Sectors 43, as these FFIs are quickly available after the data is down-linked. We created light curves from the \texttt{tica}-processed FFIs by modeling the point-spread function (PSF) of \sname and the two nearby bright sources (see Figure~\ref{fig:tpf}). While aperture photometry provided a decent light curve, we found that modeling each star (3 in total) with a 2D Gaussian created a less contaminated light curve for \sname (Figure~\ref{fig:transits}, top row).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!ht]
\begin{center}
\includegraphics[width=0.465\textwidth,trim={0.25cm 0 4.5cm 0}]{figures/folded_compare.pdf}
\caption{Phase-folded \tess\ data (gray) with the new best-fit model (color) compared to the best-fit parameters from the original \textit{K2} data (black). The residuals between the data and each fit are shown underneath. A comparison of measured $R_p/R_\star$ are presented at the bottom, with a 1-to-1 line plotted in black for reference. The transit depth from the \textit{K2} data is deeper than that from \tess, particularly for \planetb and \planete. }
  \label{fig:compare}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

where $y$ is the flux array, and $\sigma_l$  and $\sigma_j$ are used to define the light curve noise and in-transit jitter. $\sigma_l$  and $\sigma_j$ are also used as the first and second terms in our rotation model, which we defined as a stochastically-driven, damped harmonic oscillator, defined by the \texttt{SHOTerm} in \texttt{celerite} \citep{dfm17}.

@@

\section{Discussion} \label{sec:discussion}

We find transit timing variations (TTVs) for \planetknown to range from \textcolor{red}{n - n days}. A full TTV analysis using \textit{K2}, \textit{Spitzer}, and \tess\ ephemerides is presented in Livingston et al. (in prep).

\subsection{Differences in Measured Radii}

@@

The ability of a planet to retain hazes/transition hazes is positively correlated with its equilibrium temperature, $T_{eq}$, internal temperature $T_{int}$, and mass, $M_{core}$. \allplanets have calculated $T_{eq}<1000$\,K \citep{David2019a}. Young planets are believed to have high $T_{int}$. The combination of these two parameters make these planets more likely to have extended atmospheres, while outflow winds lead to the formation of transition hazes \citep{gao20}.

If \allplanets host extended haze-dominated atmospheres, they should have smaller transit radii at mid-IR wavelengths \citep{gao20}. However, without mass estimates for \allplanets, it is difficult to determine if this is the cause of the different transit depths we measured between the \textit{K2} and \tess\ data. While mass measurements are challenging for young planets due to the underlying stellar activity,  ...

There was a single transit of \planete in the original \textit{K2} data, which occurred roughly in the middle of the campaign. Since no other transits were detected, this provides a lower period limit of 36~days. Using the original transit timing from \textit{K2} and this new transit timing from \tess, we can place further constraints on the period.

We localized a 1-day window around each transit of \planete and ran a GP fit using \texttt{exoplanet} for the transit parameters, assuming a constant depth between the \textit{K2} and \tess\ observations. We then compute possible periods 
\begin{equation}
    P = \frac{1}{n} \left(T_{mid,TESS} - T_{mid, K2}\right)
\end{equation}
where $T_{mid}$ are the transit midpoints from \textit{K2} and \tess\ and $n$ is an integer. We then create a light curve using this period, and compute the log-likelihood for the given set of transit parameters. We tested discrete periods ranging from $44- 56$~days. We limit the periods used due to computation time.

@@

We find the transit depths of \allplanets as observed by \tess\ are shallower than previous \textit{K2} observations. We postulate this could be due to ongoing dusty outflows, that make the transit depth appear shallower. However, measuring the transit depths at longer wavelengths could confirm this.

@@

\begin{acknowledgments}
It is a pleasure to thank the Astronomical Data Group at the Flatiron Institute for helpful discussions. ADF acknowledges support from the National Science Foundation Graduate Research Fellowship Program under Grant No. (DGE-1746045).

@@

\software{\texttt{exoplanet} \citep{exoplanet2021},
          \texttt{EVEREST 2.0} \citep{luger18},
          \texttt{lightkurve} \citep{lightkurve},
          \texttt{matplotlib} \citep{matplotlib},
          \texttt{PyMC3} \citep{Salvatier16},
          \texttt{starry} \citep{luger19},
          \texttt{theano} \citep{theano},
          \texttt{tica} \citep{fausnaugh20},
          \texttt{batman} \citep{Kreidberg15}
          }

\bibliography{main}{}
\bibliographystyle{aasjournal}

%% Include this line if you are using the \added, \replaced, \deleted
%% commands to see a summary list of all changes at the end of the article.
%\listofchanges

>>>>

\title{V1298 Tau with TESS: Updated Ephemerides, Radii, and Period Constraints for V1298~Tau~e}

@@

\author[0000-0002-9328-5652]{Daniel Foreman-Mackey}
\affiliation{Center for Computational Astrophysics, Flatiron Institute, New York, NY 10010, USA}

@@

@@

%%%%%%%%%%%%%%%%%%%%
% abstract must be < 250 words
% currently at 180 words
%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
\sname is a young (20--30~Myr) solar analogue hosting four transiting exoplanets with sizes between $0.5 - 0.9 R_J$. Given the system's youth,  it provides a unique opportunity to understand the evolution of planetary radii in the same stellar environment. \sname was originally observed 6 years ago during K2 Campaign 4. The extended mission of NASA's Transiting Exoplanet Survey Satellite (\tess) includes observing the ecliptic plane. Here, we present new photometric observations of \sname from the 10-minute TESS Full-Frame Images. We use the TESS data to update the transit-timing for \allplanets as well as compare newly observed radii to those derived from the K2 light curve, finding shallower transits in the redder \tess\ bandpass at the $1-2\sigma$ level. We hypothesize the shallower transits observed by \tess\ may be due to high-altitude hazes in the extended atmospheres of these young planets. We additionally catch a transit of \planete, which was a single-transit event in the previous data, allowing us to further constrain the period of this farthest known planet. \sname is the target of several ongoing and future observations, including with the James Webb Space Telescope. These updated ephemerides will be crucial for accurately recovering transit events and understanding any future Doppler tomographic or transmission spectroscopy signals.\end{abstract}

@@

The youth of \sname, also known as K2-309 and [WKS96] 4, was initially established on the basis of its strong X-ray emission \citep{Wichmann1996} and high photospheric lithium abundance \citep{Wichmann2000}. A blind search for co-moving stars using Gaia DR1 astrometry data found \sname was co-moving with 8 other stars \citep[Group 29 in][]{Oh2017}. In a kinematic study of the Taurus star-forming region \citet{Luhman2018} used Gaia DR2 data to extended the membership of this group and derived an age of $\sim$~40~Myr. However, more recent analyses based on Gaia EDR3 astrometry suggests \sname may belong to either the D2 or D3 subgroups of Taurus, both of which have estimated ages $\lesssim$10~Myr \citep{gaidos21, Krolikowski2021}. Other studies focused specifically on the \sname system have estimated its age to be 23$\pm$4~Myr from comparison with empirical and theoretical isochrones \citep{David2019b}, or 28$\pm$4~Myr from isochrone fitting to the \citet{Luhman2018} Group 29 membership list given Gaia EDR3 data \citep{johnson21}. While the precise age of \sname remains uncertain, most estimates fall in the 10--40~Myr range and we adopt $t \approx$~20--30~Myr.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!ht]
\begin{center}
\includegraphics[width=0.46\textwidth,trim={0.25cm 0 0 0}]{figures/TESSaperture.pdf}
\caption{The \tess\ target pixel file (TPF) overlaid with a sky image of \sname taken with the Sloan Digital Sky Survey (SDSS) i-band. \sname is shown in the white circle; nearby sources with \tess\ magnitudes $< 14$ are marked with orange x's. There are two bight nearby sources to \sname. However, selecting the one pixel aperture (white square) encapsulates only \sname and therefore contamination from the nearby sources is negligible. \href{https://github.com/afeinstein20/v1298tau_tess/blob/main/src/figures/tpf.py}{\github}} \label{fig:tpf}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

During its Extended Mission Cycle 4, \tess\ is re-observing many of the previous \textit{K2} fields. \sname (TIC 15756231) was observed by \tess\ in Sectors 43 (UT 16 Sep 2021 to UT 12 Oct 2021) at 10-minute cadence within the Full-Frame Images (FFIs). We used the \texttt{tica} \citep{fausnaugh20} software to download calibrated FFIs each orbit for Sectors 43, as these FFIs are quickly available after the data is down-linked.

We created light curves from the \texttt{tica}-processed FFIs by modeling the point-spread function (PSF) of \sname and the two nearby bright sources (see Figure~\ref{fig:tpf}), following the PSF modeling routine in \cite{feinstein19}. In summary, we calculated and maximized the likelihood value of four parameters per each Gaussian: the $x$ and $y$ width, amplitude, and a rotational term. The Gaussian fits are allowed to vary at each time step. While aperture photometry provided a decent light curve, we found that modeling each star (3 in total) with a 2D Gaussian created a less contaminated light curve for \sname. Our extracted light curve is shown in the top row of Figure~\ref{fig:transits}.

@@

@@

where $y$ is the flux array, and $\sigma_l$  and $\sigma_j$ are used to define the light curve noise and in-transit jitter. $\sigma_l$  and $\sigma_j$ are also used as the first and second terms in our rotation model, which we defined as a stochastically-driven, damped harmonic oscillator, defined by the \texttt{SHOTerm} in \texttt{celerite2} \citep{dfm17}.

@@

There was a single transit of \planete in the original \textit{K2} data, which occurred roughly in the middle of the campaign. Since no other transits were detected, this provides a lower period limit of 36~days. Using the original transit timing from \textit{K2} and this new transit timing from \tess, we place further constraints on the period by sampling over discrete periods, similarly to \cite{becker19}. For this analysis, we used the \texttt{EVEREST 2.0} \citep{luger18} version of the \textit{K2} light curve.

We localized a 1-day window around each transit of \planete and ran a GP fit using \texttt{exoplanet} for the transit parameters, assuming a constant depth between the \textit{K2} and \tess\ observations. We assumed there is no correlation between transit parameters fit and the period we are fitting for. We then compute possible periods 
\begin{equation}\label{eq:period}
    P = \frac{1}{q} \left(T_{mid,TESS} - T_{mid, K2}\right)
\end{equation}
where $T_{mid}$ are the transit midpoints from \textit{K2} and \tess\ and $q$ is an integer. We assume a uniform prior, i.e. we have no prior preference for a specific harmonic, $q$. At each step of the sampling process, we compute a new light curve with different orbital periods, given by Equation~\ref{eq:period}. The log likelihood of each individual light curve, i.e. period, is saved, along with the sum of all log likelihoods for each period. The summation of all log likelihoods is used to compute the posterior probability for each sampled value of $q$.

We chose to test periods from $40-56$~days. We ran 3 chains with 500 tuning steps and 3000 draws. Similarly to the transit fits, we discarded the tuning steps before our analysis. We show our results in Figure~\ref{fig:period_e}, where we plot the median period for each tested harmonic against the posterior probability of each harmonic.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!ht]
\begin{center}
\includegraphics[width=0.45\textwidth,trim={0.25cm 0 0cm 0}]{figures/periode.pdf}
\caption{Our calculated log likelihood to constrain the period of \planete using transit timings from \textit{K2} and \tess. We tested discrete periods from $44-56$\,days and find the most likely period to be \textcolor{red}{n}.} \label{fig:period_e}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Differences in Measured Radii} \label{sec:radii}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!ht]
\begin{center}
\includegraphics[width=0.465\textwidth,trim={0.25cm 0 4.5cm 0}]{figures/folded_compare.pdf}
\caption{Phase-folded \tess\ data (gray) with the new best-fit model (color) compared to the best-fit parameters from the original \textit{K2} data (black). The residuals between the data and each fit are shown underneath. A comparison of measured $R_p/R_\star$ are presented at the bottom, with a 1-to-1 line plotted in black for reference. The transit depth from the \textit{K2} data is deeper than that from \tess, particularly for \planetb and \planete.} \label{fig:compare}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The ability of a planet to retain hazes/transition hazes is positively correlated with its equilibrium temperature, $T_{eq}$, internal temperature $T_{int}$, and mass, $M_{core}$. \allplanets have calculated $T_{eq}<1000$\,K, assuming an albedo = 0 \citep{David2019a}. Young planets are believed to have high $T_{int}$ due to ongoing gravitational contraction \citep{gu04}. The combination of these two parameters make these planets more likely to have extended atmospheres, while outflow winds lead to the formation of transition hazes \citep{gao20}.

If \allplanets host extended haze-dominated atmospheres, they should have smaller transit radii at mid-IR wavelengths \citep{gao20}. However, without mass estimates for \allplanets, it is difficult to determine if this is the cause of the different transit depths we measured between the \textit{K2} and \tess\ data. While radial velocity mass measurements are challenging for young planets due to the underlying stellar activity, efforts should be taken to try and constrain this parameter \citep{crockett12}.

Another, perhaps more promising approach, would be to derive the planet masses using transit-timing variations \citep[TTVs;][]{agol18}. A full TTV analysis using \textit{K2}, \textit{Spitzer}, and \tess\ ephemerides is presented in Livingston et al. (in prep).

@@

We find the transit depths of \allplanets as observed by \tess\ are shallower than previous \textit{K2} observations. We postulate this could be due to ongoing dusty outflows, that make the transit depth appear shallower. However, measuring the transit depths at longer wavelengths could corroborate this.

@@

\begin{acknowledgments}
We thank Rodrigo Luger for developing \texttt{showyourwork!} \citep{luger21} and helping us debug this letter. It is a pleasure to thank the Astronomical Data Group at the Flatiron Institute for helpful discussions. ADF acknowledges support from the National Science Foundation Graduate Research Fellowship Program under Grant No. (DGE-1746045).

@@

\software{\texttt{exoplanet} \citep{exoplanet2021},
          \texttt{EVEREST 2.0} \citep{luger18},
          \texttt{lightkurve} \citep{lightkurve},
          \texttt{matplotlib} \citep{matplotlib},
          \texttt{PyMC3} \citep{Salvatier16},
          \texttt{starry} \citep{luger19},
          \texttt{theano} \citep{theano},
          \texttt{tica} \citep{fausnaugh20},
          \texttt{batman} \citep{Kreidberg15},
          \texttt{showyourwork!} \citep{luger21},
          \texttt{astropy} \citep{astropy:2013, astropy18},
          \texttt{astroquery}\citep{astroquery19},
          numpy \citep{numpy}
          }

\bibliography{main}
\bibliographystyle{aasjournal}
